---
title: "Introduction to genefindr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to genefindr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

genefindr provides rapid gene characterization by querying eight public databases simultaneously. Instead of manually searching GeneCards, Open Targets, Human Protein Atlas, and PubMed separately, genefindr synthesizes all of this into a single function call.

## Basic usage

The main function is `findr()`. At minimum it requires a gene symbol:

```{r eval=FALSE}
library(genefindr)
findr("TP53")
```

For disease-specific context, add a `site` or `disease` argument:

```{r eval=FALSE}
findr("TP53", site = "breast")
findr("APOE", disease = "alzheimer")
```

## Supported cancer sites

The `site` argument accepts the following values:
`breast`, `prostate`, `lung`, `colon`, `ovarian`, `liver`, `brain`, `pancreatic`, `skin`, `blood`

## Multiple genes

Use `findr_multi()` to characterize several genes at once:

```{r eval=FALSE}
findr_multi(c("TP53", "BRCA1", "MYC"), site = "breast")
```

## Multi-site comparison

Compare a gene across multiple cancer types:

```{r eval=FALSE}
findr("TP53", site = c("breast", "lung", "colon"))
```

## Exporting results

Save results as a data frame or CSV:

```{r eval=FALSE}
results <- findr_multi(c("TP53", "BRCA1"), site = "breast", output = "table")
write.csv(results, "candidates.csv")
```

## Non-coding RNAs

genefindr also supports non-coding RNA genes. Protein-based fields are automatically skipped:

```{r eval=FALSE}
findr("MALAT1", site = "lung")
```

## Data sources

genefindr integrates data from eight databases:

| Database | Data provided |
|----------|--------------|
| MyGene.info | Gene name, type, summary |
| Open Targets | Disease association scores |
| Human Protein Atlas | Protein evidence, antibody availability |
| UniProt | Molecular weight, subcellular location, isoforms |
| GTEx | Normal tissue expression |
| cBioPortal/TCGA | Tumor mutation frequency |
| PubMed | Publication counts |
| ClinVar | Clinical variant counts |

## Notes

- All data sources are free and open access
- Results reflect the current state of each database at time of query
- Mutation frequency data is sourced from TCGA PanCancer Atlas 2018
- For genes with multiple isoforms, verify that your antibody targets the correct isoform
